---
title: "Analysing zotQDA exports"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysing zotQDA exports}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6,
                      fig.height = 3.5)
```

```{r setup}
library(qdaR)
```

## Where the data comes from

The plugin exports its data as a small family of CSV files. The first column
of every file says what kind of export it is and which version of the exchange
format it follows. `qdaR` checks both against the contract file and stops with
an error when it meets a version it has never seen -- guessing would be worse
than failing here, since a misread column ends up in somebody's findings
without anyone noticing.

```{r}
qda_formats()[, c("format", "file", "grain")]
```

This vignette runs entirely on reference files installed with the package, so
you can follow along without a Zotero installation:

```{r}
frag <- qda_read_fragments(qda_example("easyqda-fragments.csv"))
attr(frag, "qda_format")
attr(frag, "qda_grain")
```

## Codes carry an identity, not just a name

Codes get renamed, moved and merged while a project matures. Every export
therefore names each code twice: `code` holds the path a person reads, and
`codeId` holds a stable identifier that stays put through all of that
housekeeping. The distinction matters for any analysis that runs more than
once. If it groups by the path, a code vanishes from the results as soon as
somebody renames it in Zotero; grouped by `codeId`, it simply follows along.

```{r}
frag[, c("code", "codeId")]
```

## The plugin's figures, drawn here

```{r}
qda_plot_frequencies(frag)
```

```{r}
hist <- qda_read_history(qda_example("easyqda-history.csv"))
qda_plot_saturation(hist)
```

The original Vega-Lite charts can also be rendered unchanged with
`qda_spec_render()` when the *vegawidget* package is available -- useful when
a figure must look exactly as it did in the plugin.

## What the plugins deliberately leave out

qdaZ sticks to description and never runs a significance test. That is a
considered position, not a gap: an inferential statistic invites claims that
many qualitative designs cannot carry. If your design does support one, this
is where you run it -- and you pick the test yourself.

```{r}
res <- suppressWarnings(qda_chisq(frag, group = "citekey"))
res$table
res$cramers_v
res$expected_ok
```

Note `expected_ok`. With the tiny reference table the chi-squared
approximation does not hold, so the function reports Fisher's exact test and
says so rather than printing a p-value that looks respectable and is not.

Codes can also be arranged by the segments they share:

```{r}
qda_code_distance(frag, min_n = 1)
```

```{r}
cl <- qda_cluster(frag, min_n = 1)
cl$cophenetic
```

The cophenetic correlation is `NA` here because two codes give a single
distance, which has no variance to correlate. A dendrogram always looks
convincing; this number tells you whether it deserves to.

## Consensus data

A code-system consensus run in zotQDA ends with a mapping: this coder's code
corresponds to that consensus code. The mapping travels as ordinary data, and
`qda_apply_mapping()` does nothing more than add a `consensusCode` column next
to the original coding. Nothing gets rewritten. Had the codings themselves
been rewritten to the consensus system, any agreement you later computed on it
would come out inflated -- the disagreements would have been edited away
first.

```{r}
map <- qda_read_mapping(qda_example("easyqda-konsens-abbildung.csv"))
head(qda_apply_mapping(frag, map)[, c("code", "consensusCode")])
```

## How well did the coders agree?

Reliability is the one place where an independent reimplementation earns its
keep. qdaR recomputes the coefficients the plugin reports, in a different
language, from the exported file alone -- so a figure that appears in a
methods section has been produced twice, by two code bases that share nothing
but the contract. The package's test suite checks this against frozen
plugin results on randomly generated coder matrices.

The fragments export is long; the measures need a unit-by-coder matrix.

```{r}
frag2 <- data.frame(
  annotationKey = rep(paste0("s", 1:6), each = 2),
  codedBy = rep(c("ann", "bob"), 6),
  code = c("A", "A", "A", "A", "A", "A",
           "A", "A", "A", "A", "B", "A")
)
u <- qda_units(frag2)
qda_agreement(u)
```

Note what happened there: the coders disagreed once in six segments, and
Cohen's kappa still came out at zero. That is not a defect of the coding but
the marginals -- almost everything is `A`, so chance alone would produce this
much agreement. Gwet's AC1 is the coefficient that stays interpretable in
that situation, which is why both are reported side by side. A single
coefficient never settles the question.

Building the matrix forces two decisions. Both are easy to get silently wrong
in a hand-rolled script, so `qda_units()` makes you take them consciously:

* A segment that no coder coded still says something: both coders judged it
  irrelevant, and that is agreement too. Such segments only enter the matrix
  when you pass the `uncoded` export. Without it, your figures describe just
  the material somebody marked -- a different and usually friendlier question.
* When one coder gave a segment several codes, there is no single value to
  compare. The segment is set aside and counted in `multi_set_aside`. Quote
  that count next to the coefficient; a kappa that quietly dropped a tenth of
  the material is not the
  kappa of the study. The honest way to include such material is the
  per-code binary view, `qda_units_binary()`.

### Where in the code system the agreement is lost

A hierarchical code system can be read at several resolutions. Coders who
split over `Belastung/beruflich` against `Belastung/privat` still agree that
the segment is about `Belastung`. Flattening the paths level by level shows
where the agreement breaks down -- a statement about the code system rather
than about the coders.

```{r, fig.alt = "Agreement measures against the number of code-system levels kept"}
u2 <- cbind(ann = c("A/x", "A/y", "B/x", "B/y"),
            bob = c("A/y", "A/y", "B/x", "B/x"))
qda_level_agreement(u2)
qda_plot_level_agreement(u2)
```

And when the number is disappointing, the confusion table says which pairs of
categories cost it -- usually a handful, and usually the ones whose
definitions need work.

```{r}
qda_confusion(u2, only_disagreements = TRUE)
```
