---
title: "Figures"
---


## The plugin's figures in ggplot2

```r
qda_plot_frequencies(frag)       # how often each code was used
qda_plot_code_matrix(frag)       # codes against documents
qda_plot_timeline(history)       # codings over time, per coder
qda_plot_saturation(history)     # how the code system stopped growing
qda_plot_mds(frag)               # a map of codes by shared segments
qda_plot_level_agreement(units)  # agreement per level of the code system
```

What comes back is an ordinary `ggplot` object. Nothing is locked down, so
your usual habits apply: swap the theme, rescale an axis, compose several
panels with patchwork, write it out with `ggsave()`.

Everything that follows on this page is *live*: the code runs while the
documentation is built, and the figures are its actual output. The material
is [the example study](example-data.qmd) that ships with the package: a
synthetic study of workload in nursing, nine codes in three themes, eight
interviews from two wards, coded by two people:

```{r}
#| message: false
library(qdaR)
library(ggplot2)

frag <- qda_read_fragments(qda_example("zotqda-fragments-demo.csv"))
```

Correspondence analysis puts codes **and** interviews onto one shared map,
and it brings its own caveat with it: the share of inertia the two
dimensions show is printed on the figure itself — because a map without
that number looks just as trustworthy at 30% as at 90%. And since
`qda_ca()` returns ordinary data frames, the map is nothing more than
ggplot2:

```{r}
#| label: fig-ca
#| fig-cap: "Codes and documents on one correspondence-analysis map. The two wards separate along dimension 1, each surrounded by its typical codes. The figure states itself how much of the structure it shows."
res <- qda_ca(frag)
map <- rbind(
  data.frame(res$row_scores, name = rownames(res$row_scores), what = "code"),
  data.frame(res$col_scores, name = rownames(res$col_scores), what = "interview"))

ggplot(map, aes(V1, V2, colour = what, shape = what)) +
  geom_hline(yintercept = 0, colour = "grey85") +
  geom_vline(xintercept = 0, colour = "grey85") +
  geom_point(size = 2) +
  geom_text(aes(label = name), vjust = -0.8, size = 3, show.legend = FALSE) +
  scale_x_continuous(expand = expansion(mult = 0.15)) +
  labs(subtitle = sprintf("%.0f%% of the inertia in two dimensions",
                          100 * sum(res$inertia_share)),
       x = "dimension 1", y = "dimension 2",
       colour = NULL, shape = NULL) +
  theme_minimal()
```

Which codes sit on the same passages? The dendrogram groups codes by
co-occurrence. The cophenetic correlation in its title says how faithfully
the tree preserves the distances:

```{r}
#| label: fig-dendro
#| fig-cap: "Codes grouped by co-occurrence (Jaccard distance). The strain/coping pairs built into the material come back out as clusters."
cl <- qda_cluster(frag)
plot(cl$hclust, sub = "", xlab = "", ylab = "Jaccard distance",
     main = sprintf("Cophenetic correlation %.2f", cl$cophenetic))
```

The code-by-document matrix shows the same structure as a heatmap. One
glance is enough to see which interviews carry which themes:

```{r}
#| label: fig-matrix
#| fig-cap: "Codings per code and interview. The block structure between the two wards is the group difference the map above projects."
qda_plot_code_matrix(frag)
```

## Reading the numbers off a figure

Behind each plot sits a function that hands you the same numbers as a data
frame. Supplements and reviewers tend to want those rather than the picture:

```{r}
head(qda_code_counts(frag))
```

```r
qda_code_matrix(frag, long = FALSE) # codes x documents
qda_timeline(history)               # cumulative codings per coder
qda_saturation(history)             # distinct codes over time
```

One thing to know about `qda_timeline()`: it counts `add` events and ignores
`remove`. A coding somebody applied and then withdrew did happen, but it is not
part of what you ended up with, and a growth curve that keeps counting it
flatters the result.

## Figures that carry their own caveat

Two of these figures come with a number attached: `qda_mds()` returns
`goodness`, the share of eigenvalue mass its two dimensions carry, and
`qda_cluster()` returns `cophenetic`. Neither number is decoration. A code map
showing a third of the structure and one showing nearly all of it look exactly
alike, and so do a trustworthy dendrogram and a misleading one.

## The plugin's own charts, unchanged

Every chart in qdaZ can be saved as a data-plus-specification pair. Rendering
the specification reproduces the figure exactly as it looked in Zotero:

```r
spec <- qda_spec_read("frequencies.json")
qda_spec_render(spec, qda_code_counts(frag))
```

This needs the *vegawidget* package, which is suggested rather than required —
`install.packages("vegawidget")`. Use it when a figure must match the plugin;
use the `qda_plot_*` functions when it must match the rest of your manuscript.

## The same figures in Python

[qdaPy](https://qdapy.zotqda.org/) draws all of these with plotnine and
seaborn from the same tables. If your project mixes languages, the figures
will still agree.
